---
title: "PDF Example"
author: "Jason Thomas"
date: "1/7/2020"
output: 
  pdf_document:
    latex_engine: xelatex
linkcolor: red
bibliography: vaLitReview.bib
geometry: margin = 1in
header-includes:
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage{xcolor}
---

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = TRUE)
read_chunk("rmarkdown_advanced.R")
```


First some citations from a paper.  @mahapatra07 discuss ways to assess
the importance of CRVS.  Demographers should love CRVS for several reasons [@mahapatra07].  Mahpatra ([-@mahapatra07])and
colleagues have discussed these points.

* include a cache example

## Generate some data

```{r generate_data}
```

An example of in-line code: the third observation in x is `r x[3]`.

## Regression Model

```{r model_1}
```


## Tables
```{r table}
library(kableExtra)
dt <- mtcars[1:5, 1:6]
kable(dt, "latex", booktabs = T)%>%kable_styling(latex_options = "striped")
```

## More Examples with tables

* [Manual for kableExtra](https://haozhu233.github.io/kableExtra/awesome_table_in_pdf.pdf)

* Another useful R package is [`stargazer`](https://cran.r-project.org/web/packages/stargazer/vignettes/stargazer.pdf)

## Stata, I miss you!

```{stata, engine.path = "/Applications/Stata/StataMP.app/Contents/MacOS/stata-mp", error = TRUE, highlight.output = TRUE}
sysuse auto
su price
```

```{stata, engine.path = "/Applications/Stata/StataMP.app/Contents/MacOS/stata-mp", error = TRUE, highlight.output = TRUE}
sysuse auto
scatter weight mpg
graph export "stata_plot.pdf", replace
```

![]("./stata_plot")

# References
