---
title: R Markdown Example to Word
subtitle: (and other stuff)
author:
  - Jason Thomas^[Inst. Pop. Research, OSU]
  - Leo Messi^[Barca, Argentinian Natl. Team]
date: "`r Sys.Date()`"
output:
  word_document:
    toc: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(furniture)
library(flextable)
library(knitr)
library(dplyr)
data("airquality")
data(mtcars)
fig_count <- 1
```

## Intro

Here is an example of `R markdown` to MS Word.

## flextable

```{r flex1}
ft <- flextable(airquality[ sample.int(10),])
ft <- add_header_row(ft,
  colwidths = c(4, 2),
  values = c("Air quality", "Time")
)
ft <- theme_vanilla(ft)
ft <- add_footer_lines(ft, "Daily air quality measurements in New York, May to September 1973.")
ft <- color(ft, part = "footer", color = "#666666")
ft <- set_caption(ft, caption = "New York Air Quality Measurements")
ft
```

```{r flex2}
myft <- flextable(head(mtcars), 
  col_keys = c("am", "carb", "gear", "mpg", "drat" )) %>%
  italic(j = 3) %>%
  color(~ drat > 3.5, ~ drat, color = "red") %>%
  bold(~ drat > 3.5, ~ drat, bold = TRUE) %>%
  add_header_row(values = c("some measures", "other measures"),
  colwidths = c(3, 2)) %>%
  align(i = 1, part = "header", align = "center") %>%
  theme_zebra()
myft
```


```{r flex2_1}
myft <- flextable(head(mtcars), 
  col_keys = c("am", "carb", "gear", "mpg", "drat" ))
myft <- italic(myft, j = 3)
myft <- color(myft, ~ drat > 3.5, ~ drat, color = "red")
myft <- bold(myft, ~ drat > 3.5, ~ drat, bold = TRUE)
myft <- add_header_row(
  x = myft, values = c("some measures", "other measures"),
  colwidths = c(3, 2))
myft <- align(myft, i = 1, part = "header", align = "center")
myft <- theme_vanilla(myft)
myft
```

```{r flex3}
mod1 <- lm(mpg ~ hp + cyl, data=mtcars)
mod2 <- lm(mpg ~ hp + cyl + gear, data=mtcars)
mod3 <- lm(mpg ~ hp + cyl + gear + disp, data=mtcars)
model_names <- c("mod1", "mod2", "mod3")
models <- matrix(NA, nrow=length(coef(mod3)), ncol=length(model_names))
for (i in 1:length(model_names)) { # i = 1
  tmp_mod <- get(model_names[i])
  tmp_coefs <- coef(tmp_mod)
  models[1:length(tmp_coefs), i] <- round(coef(tmp_mod), 2)
}
model_labels <- c("Model 1", "Model 2", "Model 3")
df_models <- as.data.frame(models)
names(df_models) <- model_labels
df_models$Covariates <- names(coef(mod3))
df_models <- df_models[, c(4, 1:3)]
myft_reg <- flextable(df_models) %>%
  theme_tron()
myft_reg
```

## Table with Furniture

```{r tableF}
table1(mtcars, type = c("simple", "condensed"), output = "pandoc")
```

```{r descriptives furniture ex3}
vNames <- c("Miles per gallon", "Number of cylinders",
            "Gross horsepower")
table1(mtcars, mpg, cyl, hp, var_names = vNames,
       splitby = ~am,
       caption = "**Means and SD for subset of variables**",
       type = c("simple", "condensed"),
       output = "pandoc")
```

```{r cross tab tableX}
crossTab <- tableX(mtcars, cyl, am, type = "count")
kable(crossTab)
```

# Plot with Reference

```{r label_fig1, include=FALSE}
fig_lab_mpg_hp <- fig_count
fig_count <- fig_count + 1
```
This section has an example of including a plot.  More specifically Figure `r fig_lab_mpg_hp`.  But I think the officer package might have a better solution.

```{r cars, fig.cap = paste0("Figure ", fig_lab_mpg_hp, ": MPG vs HP")}
plot(mpg ~ hp, data = mtcars)
```

# Equations

$\hat{y_i} = \alpha + \beta_1 * X_{i,1} + \beta_2 * X_{i,2} + \epsilon_i$

and check out

* [equationmatic](https://bookdown.org/yihui/rmarkdown-cookbook/equatiomatic.html)

# YAML & other stuff

* [author affilation](https://bookdown.org/yihui/rmarkdown-cookbook/multiple-authors.html)
* [diagrams](https://bookdown.org/yihui/rmarkdown-cookbook/diagrams.html)
* officeverse
