---
title: "Slides with Revealjs"
subtitle: "R Markdown Example"
author: "Bandit Healer"
date: "`r format(Sys.Date(), format='%A, %B %d %Y')`"
output:
  revealjs::revealjs_presentation:
    transition: fade
    background_transition: concave
    reveal_options:
      slideNumber: true
      previewLinks: true
---

## R Markdown

This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

. . .

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.


## Slide with Bullets

- Bullet 1
- Bullet 2
- Bullet 3

## Slide with R Code and Output

Let's see a summary of the `mtcars` data.  Note how the output doesn't fit on the slide.

. . .

```{r}
summary(mtcars)
```

## Slide with Plot

PLOT!

```{r, echo=FALSE, fig.align='right'}
plot(cars)
```


## Another Slide with Plot

small plot

```{r, echo=FALSE, fig.height=3, fig.width=3}
plot(cars)
```


## Slide in, Fade out {data-transition="slide-in fade-out"}


## Convex in, Concave out {data-transition="convex-in concave-out"}

# Inlude Some Horizontal Slides

## Horizontal 1

## Horizontal 2

# Include a 2nd Set of Horizontal Slides


## Leaflet


## CSS color background {data-background=#ff0000}

## Full size image background {data-background="img/IPR-Vert.png"}

## Video background {data-background-video="img/background.mp4"}

## A background page {data-background-iframe="https://buckipr.github.io/R_Working_Group"}

