---
title: "Introduction to R"
author: "R Working Group"
date: "2025-01-24"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Two for the Price of One!

Buckle your seat belt because we are trying something *NEW* for this introduction
to R.  Your brain is probably saying

"This does not look like the normal script file!"

and your brain is correct!  Welcome to the world of R Markdown!  Ready for a
magic trick?  See that little ball of yarn & knitting needle (next to the word
Knit)?  Click it and **behold**!

<br>

<br>

<br>


## Goals

* Load Data into R

* Create New Variables

* Exploratory Data Analysis

* Tables & Plots?

<br>

<br>

<br>


## R Markdown (quick tour)

### Markdown

Not a lot here (we've already seen most of the magic).  For more check out:
[markdown link](https://markdownlivepreview.com/)

### R Markdown

Code chunk options:

* `echo=FALSE` -- display the R code
* `include=TRUE` -- include the results produced by running the R code

**Exercises**

* Shortcuts
  + What is the keyboard shortcut for inserting a code chunk?
  + Is there a keyboard shortcut for running all the code in a chunk?
* How do you include?
  + Why would you ever want to do this?
* Can you make a nested list with spaces between all of the items?
  + How many levels can you create in a nested list?
* Code Chunks
  + What does the code chunk option `eval=FALSE` do?
  + How do you keep warning messages from being included in your output?
  + Include a link to a good resource
* What the heck is YAML?  Find 3 super cool things you can do with it...
  1. thing 1:
  1. thing 2:
  1. thing 3:
* Markdown
  + How do you add a page break?
  + Can you include a footnote in your doc?  an endnote?  a link
  to another section in your doc?!?
  + *Can you include citations/references?!?*
* Anything else???

<br>

<br>

<br>


## Loading Data into R (and kicking it out)

Let's load a practice data set from the web, then do some stuff.

```{r gradsch_data}
# comment: Load example data set 
gradsch_data <- read.csv("https://stats.idre.ucla.edu/stat/data/binary.csv")
gradsch_data
```

As you will learn, Jason is fairly lazy and thus the commentary will now cease
since his poor fingers are likely too be quite fatigued by this point! (sob, sob)


```{r take_2}
head(gradsch_data) ## much better
## dim() is also useful == (nrow(), ncol())
```


```{r take_3}
summary(gradsch_data) ## now you are just showing off!
```

```{r admit}
table(gradsch_data$admit)
```

#### Ignore for now

(but run the chunks!  we'll use the results later)

```{r interestify}
gradsch_data$admit[c(11, 22, 33, 44, 55, 66)] <- NA
gradsch_data$gre[c(55, 66, 77, 88, 99, 111)] <- NA
gradsch_data$gre[193] <- 29875
gradsch_data$gpa[c(122, 133, 144, 155, 166, 177, 188)] <- NA
gradsch_data$rank[c(11, 22, 77, 99, 122, 133, 177, 199, 222, 333)] <- NA
```

#### Outro

```{r outro}
getwd()  ## this is where the file will be save (change with setwd("path"))
dir()    ## print the contents of the current directory
write.csv(gradsch_data, file = "gs.csv")
```

**Exercises**

* Change the next (sub)item so that it includes the actual number of observations
(instead of "a lot").
  + There are N = `r print("a lot")` observations in this data set.
* What does the function `str` do?
* How do you exprot data to a CSV file so that first column or row numbers is
not included?
* Read in the data set `gd.csv`
  + Print out the first observation for rank
  + Print out the third observation all variables
  + Create a new data set that only includes the first 4 observations
* Anything else???

<br>

<br>

<br>


## Create New Variables

```{r new_vars}
gradsch_data$top_rank <- NA
table(gradsch_data$top_rank, gradsch_data$rank, useNA = "always")
```

```{r top_rank}
gradsch_data$rank == 4
gradsch_data$top_rank[ gradsch_data$rank == 4 ] <- 1
# now assign 0 if rank == 1, 2, or 3
# make a table to check your coding
```


```{r super_star}
summary(gradsch_data$gre)
summary(gradsch_data$gpa)

# super_star = 1 if above mean gre AND gpa
gradsch_data$ss <- 0
gradsch_data$gre > mean(gradsch_data$gpa) # can you fix this? (2 problems)

# table(gradsch_data$gre > mean(gradsch_data$gpa),
#       gradsch_data$gpa > mean(gradsch_data$gpa),
#       useNA = "always")
# indicator <- gradsch_data$gre > mean(gradsch_data$gpa) &
#       gradsch_data$gre > mean(gradsch_data$gpa)
## use our indicator to finish coding super_star
## what happened to the NA's?
```

**Exercises**

* What does the function `is.na` do?  Use this function to create a new variable
`ss2` where all the missing value are coded to 0.
* "Or" is implemented in R with a vertical bar |.  Create a new variable called
star that takes a value of 1 if the student's gre is above the 3rd quantile OR
the students gpa is above the 3rd quantile (and 0 otherwise).
* Create a new data frame that only includes complete cases (i.e. no missing data)
* Anything else???

<br>

<br>

<br>


## Exploratory Data Analysis


**Exercises**

* What are the 5 most useful functions in `dplyr`?
  + Describe what each function does and take it for a spin
  + What is pipe? (in the R language)  Is it useful?
* Anything else???

<br>

<br>

<br>


## Tables & Plots

**Exercises**

* The function `lm()` will run a linear regression.
  + Regress admit on gpa (model 1).
  + Regress admit on gre (model 2).
  + Regress admit on gpa + gre (model 3).
  + Regress admit on gpa + gre + gpa-gre interaction (model 4).
  + Make a table of your regression results (coefficients and standard errors).
  + Write the table to a CSV file (that would make an journal editor proud).
* Guess what the function `()` does.
  + Plot gre vs. gpa.
  + Add a red lines that indicate the mean of gpa and the mean of gre.
  + If the student was admitted, change the plotting character to "A",
  and if the student was not addmitted, change the plotting character to "N"
  + Repeat these steps with `ggplot2`
* Anything else???
