---
title: |
  | *R Working Group*
  |
  | **Data Management in**  \faRProject
  | **using `dplyr`**
author: |
  | Jason Thomas
  | thomas.3912@osu.edu
date: |
  | Friday, February 17th, 2023
  | 11am - 12:30pm
header-includes:
  - \usepackage{xcolor}
  - \usepackage{fontawesome5}
output: 
  beamer_presentation:
    latex_engine: xelatex
    toc: FALSE
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```
\newcommand{\R}{\color{blue} \faRProject \; \color{black}}

## Motivation

* \R -- past 5 years (or so) has included great (user-friendly) strides in data management

    + [`dplyr` package](https://dplyr.tidyverse.org/)  -- "grammar of data manipulation,
    providing a consistent set of verbs that help you solve the most common data manipulation
    challenges"

    + a lot of functions coded in C++ (for speed)

    + many other packages make use of `dplyr`

* Goal for today is to establish basic fluency with this grammar


## Background

* `dplyr` is part of [`tidyverse`](https://www.tidyverse.org/)

  + `ggplot2`, `forcats`, `tibble`, `readr`, `stringr`,  `tidyr`, `purrr`
  + may also want to check out `tidycensus`[https://walker-data.com/tidycensus/articles/basic-usage.html]

* `dplyr` logic: "By constraining your options, it helps you think about your data manipulation challenges."

  + 5 commands will take you a long way
  + readability and simplifying code (with pipes)

* \R   {\bf for Data Science}

  + [https://r4ds.had.co.nz/](https://r4ds.had.co.nz/)

```{r, dplyr, eval=FALSE}
library(dplyr)
```


## Grammar

* **Rows** -- selecting and organizing observations/cases

    + **Groups of Rows**

* **Columns** -- cleaning and creating new variables

* **Merging Data** (not going to go into this today)

* Additional features of the language
    + tibble data structure (similar to data frames, but crankier)
    + "pipe" %>%  for stringing multiple commands together
    + useful for keeping number of objects to a minimum and for
    plotting (e.g., adding separate symbols for subgroups)


## Grammar for Rows

* `filter()`

* `slice()`

* `arrange()`

* *Groups of rows*: `summarize()` 
  + useful companion: group_by()
  + collapse across rows with `summarize()`


## Grammar for Columns

* `select()`

* `rename()`

* `mutate()`

