---
title: Introduction to GitHub <br><br>
  and Integration with <br><br>
  RStudio <br><br>
author: "Jason Thomas"
institute: "R Working Group"
date: "Oct. 2nd, 2025"
output:
  xaringan::moon_reader:
    css: ["style_rgroup.css"]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
library(knitr)
library(emoji)
knitr::opts_chunk$set(tidy=TRUE, prompt=TRUE, error=TRUE)
```


# Motivation

* Messy projects, messy code, and sharing/collaboration?

* GitHub to the rescue

  + GitHub is an on-line service that stores files, tracks versions,
  and more!

* `git` is the program used for *version control* (i.e.,
managing and tracking different versions of your files)
  
  + RStudio adds features for running git on your computer
  <br> (RStudio's VC [user guide](https://docs.posit.co/ide/user/ide/guide/tools/version-control.html))
  
  + (there are other programs for version control, e.g. `subversion`)

---
# Motivation (cont.)
  
* "and more" you say??

  + renders *markdown* code (used to format text)
  <br> test drive: [https://markdownlivepreview.com](https://markdownlivepreview.com/)
  
  + websites, e.g. 
  [https://buckipr.github.io/R_Working_Group/](https://buckipr.github.io/R_Working_Group/)
  
  + create different *branches* of you project
  
  + advanced features for software developers (making it popular
  and a common source for latest versions)


---
# Getting Started

* GitHub Account

  + (more details to come)

* Install `git`

  + Windows: available for download from: [https://git-scm.com/downloads](https://git-scm.com/downloads)
  
  + MacOS: install Xcode from the App Store (includes git)

  + an excellent resource: [git book](https://git-scm.com/book/en/v2)
  
  + (if you start using git/GitHub
a lot it is well work taking a week or two and groing through the
first few chapters)


---
class: slide-font-25
# Getting Started (cont.)

* Create **repository** on GitHub 

  + a *repo* is a folder where all your project files are stored
  <br> (including those needed for managing version control)
  
  + can be on your computer (local) or on GitHub (remote)
  
  + the idea is to run git to track changes, then sync the local
  and remote repos (so your changes are available on GitHub)

* Create project in RStudio

  + This will create a new file `project_name.Rproj` with information
  that RStudio uses to set up your project (e.g., GitHub repo details)
  
  + Configure your (local) copy of the repo with the remote copy
  (on Git Hub)


---
class: inverse, center, middle

# GitHub & Git


---
## **GitHub**: signing up

Sign up for a free GitHub account:

* [https://github.com/join](https://github.com/join)

    + you will need to create a new username and password
    
    + the *Free* subscription will work just fine
    
    + there is also an 
    [educational account](https://help.github.com/en/articles/about-github-education-for-educators-and-researchers)
    that provides more resources
    
    + a verification email will be sent (check spam folder)


---
## GitHub: concepts and actions

* **Clone** - copy a repository to your computer for the first time

* **Pull** - update your local copy of the repo with any new
changes on the remote repo

  + clone is basically a pull for the first time
  
  + you can also *fetch* - download changes (so you can inspect them)
  but do not sync them with your local repo

* Make new files and changes to your local repo, then **STAGE** all
the files you want to keep track of

  + this includes adding, renaming/moving, & removing files with
  any changes that you want to track


---
## GitHub: concepts and actions (cont.)

* **Commit** - save your changes (they are now tracked)

  + typically include a message describing the commit
  
  + this creates a new version (or snap shot) of your project
  
  + you can go back and look at previous commits (or revert to that
  version) and look at differences between commits

* **Push** - send your changes to GitHub

  + local repo will now be identical to remote repo


---
## Configure git & RStudio

* GitHub requires more security than just a simple
password authentication &rarr; **Personal Access Token**

* The `usethis` package will help set this up, so let's install the
package (if we haven't already) and attach it

```{r eval=FALSE}
install.package("usethis")  # only need to run this once
library(usethis)
```

* Next we can check out the status of our `git` setup with

```{r eval=FALSE}
git_sitrep()
```


---
## Configure RStudio (cont.)

* Set our user name and email:

```{r eval=FALSE}
use_git_config(user.name = "<your name>", user.email = "<your email>")
git_sitrep() # check it worked
```

* (optional) set default branch name (same as what GitHub uses)

```{r eval=FALSE}
git_default_branch_configure()  # default is main
git_sitrep()
```

* Set Personal Access Token

```{r eval=FALSE}
create_github_token()
# opens up GitHub page in web browser
# set note, expiration, then copy the token -> Generate Token
gitcreds::gitcreds_set()
# after running the above line, paste in the PAT
```



---
## Create RStudio Project

* Change to your project folder and set it as the working directory

* Create an RStudio Project for you project folder:

```{r eval=FALSE}
create_project(".")
# the "." means use your working directory as the project folder
```

* RStudio will restart and open your new Project


---
## Create Repository

* Turn local directory into a repo configured with your GitHub settings

```{r eval=FALSE}
library(usethis)  # need to re-attach after restart
use_git()         # turns our project folder into a repository that can
                  # be synced with GitHub
```

* `use_get()` will make a few changes to the .gitignore file and ask if
you want to commit it and the .Rproj (RStudio project configuration
file) to your repository.

* RStudio will restart again and open your project folder


---
## Sync with GitHub

* Now sync the local version of your repository with GitHub

  - (this will create the repository on GitHub and include the
  .gitignore & .Rproj files if you said 'yes' to the previous step)

```{r eval=FALSE}
library(usethis)
use_github()
```
